---
title: "Final Project"
author: "Claudia Lo"
date: '2022-05-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Predicting Diabetes Readmission
This diabetes dataset from https://archive.ics.uci.edu/ml/datasets/diabetes+130-us+hospitals+for+years+1999-2008 is a record of 10 years(1999-2008) of clinical care at 130 US hospitals. Through my study, I will explore which features have a high chance of affecting readmission and predict the probability of patient readmission.

```{r}
#set working directory
getwd()
setwd("C:/Users/claud/OneDrive/Desktop/MSDHI/spr 2022/HS 614 data analytics/final project")
```
```{r}
#Import datasets
readmission<-read.csv("./data/diabetic_data.csv",na.string="?",header = TRUE, sep = ",")
```

```{r}
readmission_id<-read.csv("./data/IDs_mapping.csv",na.string="?",header = TRUE, sep = ",")
readmission_id
```

```{r}
head(readmission)
```

```{r}
str(readmission)
```

```{r}
summary(readmission)
```

```{r}
nrow(readmission)
ncol(readmission)
```

```{r}
#count number of missing values
colSums(is.na(readmission))
```
```{r}
#change chr variables to factor
readmission[sapply(readmission, is.character)] <- lapply(readmission[sapply(readmission, is.character)],as.factor)
```

```{r}
str(readmission)
```

##Cleaning data
```{r}
#removing columns that are not useful and have a lot of missing values
readmission <- readmission[, -c(1, 2, 6,11,12)]
```

```{r}
nrow(readmission)
ncol(readmission)
```

```{r}
table(readmission$max_glu_serum)
```

```{r}
table(readmission$A1Cresult)
```

```{r}
readmission <- readmission[!readmission$max_glu_serum == "None", ]
readmission <- readmission[!readmission$A1Cresult == "None", ]
```

```{r}
nrow(readmission)
```
```{r}
table(readmission$max_glu_serum)
table(readmission$A1Cresult)
```
```{r}
nrow(readmission)
ncol(readmission)
```

##Classifying outcome variable into two groups
```{r}
table(readmission$readmitted)
```
```{r}
#binary responses function
binary = function(x){
    if(x == '<30'| x == '>30'){
        return('YES')
    }else{ return('NO')}
}
```

```{r}
binary('<30')
```
```{r}
binary('>30')
```
```{r}
binary('NO')
```
```{r}
readmission$readmitted <- sapply(readmission[,c('readmitted')], function(x) binary(x))
```

```{r}
table(readmission$readmitted)
```
##EDA
```{r}
#barplot of outcome variable
library(dplyr)
library(ggplot2)
df <- readmission %>%
          filter(!is.na(readmitted)) %>% 
          group_by(readmitted) %>%
  summarise(count_m= n())
        p <-ggplot(df, aes(x = readmitted, y = count_m)) +
          geom_bar(fill = "#3DDAEB", stat = "identity") +
          geom_text(aes(label = count_m), vjust = -0.25,size = 4) + 
          theme(axis.text = element_text(colour = "blue"),text = element_text(size=10),axis.line = element_line(size = 4, colour = "grey80"),axis.text.x = element_text(angle = 90, vjust = 0.45, hjust=1)) + ylab("Number of readmissions") + xlab("Readmission") + ggtitle(labs(title = "Distribution of readmission"))
p
```

```{r}
#grouped barplot of gender and readmission
d4 <- readmission %>%
  group_by(readmitted,gender) %>%
  summarise(count=n()) %>%
  mutate(perc=count/sum(count))
bar4 <- ggplot(data=d4, aes(x=readmitted, y=perc, fill=gender)) +
  geom_bar(position = 'dodge', stat ='identity')+scale_y_continuous(labels = scales::percent)+labs(x = "readmission", y = "Percentage of readmissions", fill = "gender") + ggtitle(labs(title = "Readmissions by gender"))+ theme(text = element_text(size=10))+ geom_text(aes(label = scales::percent(perc),y = perc),position=position_dodge(width=0.7), vjust = -0.25,size=3.5)
bar4  + facet_grid(. ~ readmitted,scales='free')
```
```{r}
#grouped barplot of race and readmission
d5 <- readmission %>%
  group_by(readmitted,race) %>%
  summarise(count=n()) %>%
  mutate(perc=count/sum(count))
bar4 <- ggplot(data=d5, aes(x=readmitted, y=perc, fill=race)) +
  geom_bar(position = 'dodge', stat ='identity')+scale_y_continuous(labels = scales::percent)+labs(x = "readmission", y = "Percentage of readmissions", fill = "race") + ggtitle(labs(title = "Readmissions by race"))+ theme(text = element_text(size=10))+ geom_text(aes(label = scales::percent(perc),y = perc),position=position_dodge(width=0.7), vjust = -0.25,size=3.5)
bar4  + facet_grid(. ~ readmitted,scales='free')
```

```{r}
#density plot of time in hospital and age
Densityplot <- ggplot(data = readmission,aes(x=time_in_hospital,fill= age ))+ geom_density(alpha = 0.9,cex=1)+ggtitle(labs(title = "Distribution of time in hospital by Age"))+theme(text = element_text(size=10))
Densityplot
```

```{r}
#grouped histogram of number of inpatient visits and age
ggplot(readmission, aes(x = number_inpatient, fill = age)) + geom_histogram(colour = "black",binwidth = .5, lwd = .75, linetype= 1,position = "identity", alpha = 4) + labs(x = "number of inpatient visits", y = "Frequency", title = "Distribution of inpatient visits by Age")
```

```{r}
#grouped barplot of age and readmission
d6 <- readmission %>%
  group_by(readmitted,age) %>%
  summarise(count=n()) %>%
  mutate(perc=count/sum(count))
bar4 <- ggplot(data=d6, aes(x=readmitted, y=perc, fill=age)) +
  geom_bar(position = 'dodge', stat ='identity')+scale_y_continuous(labels = scales::percent)+labs(x = "readmission", y = "Percentage of readmissions", fill = "age") + ggtitle(labs(title = "Readmissions by age"))+ theme(text = element_text(size=10))+ geom_text(aes(label = scales::percent(perc),y = perc),position=position_dodge(width=0.7), vjust = -0.25,size=3.5)
bar4  + facet_grid(. ~ readmitted,scales='free')
```

```{r}
#density plot of number of diagnoses and readmission
Densityplot <- ggplot(data = readmission,aes(x=number_diagnoses,fill= readmitted ))+ geom_density(alpha = 0.9,cex=1)+ggtitle(labs(title = "Distribution of number of diagnoses by readmission"))+theme(text = element_text(size=10))
Densityplot
```

```{r}
class(readmission$discharge_disposition_id)
class(readmission$admission_type_id)
class(readmission$admission_source_id)
```

```{r}
readmission$discharge_disposition_id <- as.factor(readmission$discharge_disposition_id)
readmission$admission_type_id <- as.factor(readmission$admission_type_id)
readmission$admission_source_id <- as.factor(readmission$admission_source_id)
```

```{r}
str(readmission)
```

```{r}
#dataset with numeric variables
library(dplyr)
readmission_num<-select_if(readmission,is.numeric)
readmission_num
```

```{r}
#correlation
library(GGally)
ggpairs(readmission_num)
```

#Creating dummy variables
```{r}
#create dummy variables
library(caret)
d <- predict(dummyVars(~discharge_disposition_id + admission_type_id + max_glu_serum + A1Cresult + admission_source_id, data = readmission), newdata = readmission)
```

```{r}
#dataset of dummy variables
head(d)
```

```{r}
#checking column names
colnames(d)
```
```{r}
#renaming column names
colnames(d)[colnames(d) == "max_glu_serum.>200"] <- "max_glu_serum.200"
colnames(d)[colnames(d) == "max_glu_serum.>300"] <- "max_glu_serum.300" 
colnames(d)[colnames(d) == "A1Cresult.>7"] <- "A1Cresult7"
colnames(d)[colnames(d) == "A1Cresult.>8"] <- "A1Cresult8" 
```

```{r}
colnames(d)
```

##Creating diagnosis variables for diag_1, diag_2, diag_3
```{r}
myData_d<-readmission
```

```{r}
myData_d$diag_circ <- 0
myData_d$diag_resp <- 0
myData_d$diag_dig <- 0
myData_d$diag_diab <- 0
myData_d$diag_inj <- 0
myData_d$diag_musc <-0
myData_d$diag_geni <- 0
myData_d$diag_neop <-0
myData_d$diag_other <- 0
```

```{r}
#create diagnosis variable for circulatory codes:390-459, 785 
myData_d$diag_circ[(as.character( myData_d$diag_1) >= "390" & as.character( myData_d$diag_1) <= "459" |  as.character( myData_d$diag_1 ) =="785")
                   | (as.character( myData_d$diag_2) >= "390" & as.character( myData_d$diag_2) <= "459" |  as.character( myData_d$diag_2 ) == "785")
                   | (as.character( myData_d$diag_3) >= "390" & as.character( myData_d$diag_3) <= "459" |  as.character( myData_d$diag_3 ) == "785")] <- 1
#create diagnosis variable for respiratory codes: 460-519, 786
myData_d$diag_resp[(as.character( myData_d$diag_1) >= "460" & as.character( myData_d$diag_1) <= "519" |  as.character( myData_d$diag_1 ) == "786")
                   | (as.character( myData_d$diag_2) >= "460" & as.character( myData_d$diag_2) <= "519" |  as.character( myData_d$diag_2 ) == "786")
                   | (as.character( myData_d$diag_3) >= "460" & as.character( myData_d$diag_3) <= "519" |  as.character( myData_d$diag_3 ) == "786")] <- 1
#create diagnosis variable for digestive codes: 520-579, 787
myData_d$diag_dig[(as.character( myData_d$diag_1) >= "520" & as.character( myData_d$diag_1) <= "579" |  as.character( myData_d$diag_1 ) == "787")
                  | (as.character( myData_d$diag_2) >= "520" & as.character( myData_d$diag_2) <= "579" |  as.character( myData_d$diag_2 ) == "787")
                  | (as.character( myData_d$diag_3) >= "520" & as.character( myData_d$diag_3) <= "579" |  as.character( myData_d$diag_3 ) == "787")] <- 1
#create diagnosis variable for diabetes mellitus codes: 250.xx
myData_d$diag_diab[(as.character(myData_d$diag_1) > "249" & as.character(myData_d$diag_1) < "251")
                   | (as.character(myData_d$diag_2) > "249" & as.character(myData_d$diag_2) < "251")
                   | (as.character(myData_d$diag_3) > "249" & as.character(myData_d$diag_3) < "251")] <- 1
#create diagnosis variable for injury codes: 800-999
myData_d$diag_inj[(as.character( myData_d$diag_1) >= "800" & as.character( myData_d$diag_1) <= "999")
                  | (as.character( myData_d$diag_2) >= "800" & as.character( myData_d$diag_2) <= "999")
                  | (as.character( myData_d$diag_3) >= "800" & as.character( myData_d$diag_3) <= "999")] <- 1
#create diagnosis variable for musculoskeletal codes: 710-739
myData_d$diag_musc[(as.character( myData_d$diag_1) >= "710" & as.character( myData_d$diag_1) <= "739")
                   | (as.character( myData_d$diag_2) >= "710" & as.character( myData_d$diag_2) <= "739")
                   | (as.character( myData_d$diag_3) >= "710" & as.character( myData_d$diag_3) <= "739")] <- 1
#create diagnosis variable for genitourinary codes: 580-629, 788
myData_d$diag_geni[(as.character( myData_d$diag_1) >= "580" & as.character( myData_d$diag_1) <= "629" |  as.character( myData_d$diag_1 ) == "788")
                   | (as.character( myData_d$diag_2) >= "580" & as.character( myData_d$diag_2) <= "629" |  as.character( myData_d$diag_2 ) == "788")
                   | (as.character( myData_d$diag_3) >= "580" & as.character( myData_d$diag_3) <= "629" |  as.character( myData_d$diag_3 ) == "788")] <- 1
#create diagnosis variable for neoplasms codes: 140-239
myData_d$diag_neop[(as.character( myData_d$diag_1) >= "140" & as.character( myData_d$diag_1) <= "239")
                   | (as.character( myData_d$diag_2) >= "140" & as.character( myData_d$diag_2) <= "239")
                   | (as.character( myData_d$diag_3) >= "140" & as.character( myData_d$diag_3) <= "239")] <- 1
#create diagnosis variable for other diseases
myData_d$diag_other[(as.character( myData_d$diag_1) == "780") | (as.character( myData_d$diag_1) == "781")
                    | (as.character( myData_d$diag_1) == "784") | (as.character( myData_d$diag_1) >= "790" & as.character( myData_d$diag_1) <= "799")
                    | (as.character( myData_d$diag_1) >= "240" & as.character( myData_d$diag_1) <= "249") | (as.character( myData_d$diag_1) >= "251" & as.character( myData_d$diag_1) <= "279")
                    | (as.character( myData_d$diag_1) >= "680" & as.character( myData_d$diag_1) <= "709") | (as.character( myData_d$diag_1) == "782") 
                    | (as.character( myData_d$diag_1) >= "001" & as.character( myData_d$diag_1) <= "139") | (as.character( myData_d$diag_1) >= "290" & as.character( myData_d$diag_1) <= "319")
                    | (as.character( myData_d$diag_1) >= "280" & as.character( myData_d$diag_1) <= "289") | (as.character( myData_d$diag_1) >= "320" & as.character( myData_d$diag_1) <= "359")
                    | (as.character( myData_d$diag_1) >= "630" & as.character( myData_d$diag_1) <= "679") | (as.character( myData_d$diag_1) >= "360" & as.character( myData_d$diag_1) <= "389")
                    | (as.character( myData_d$diag_1) >= "740" & as.character( myData_d$diag_1) <= "759")
                    | (startsWith(as.character( myData_d$diag_1), 'E'))
                    | (startsWith(as.character( myData_d$diag_1), 'V'))
                    | (as.character( myData_d$diag_2) == "780") | (as.character( myData_d$diag_2) == "781")
                    | (as.character( myData_d$diag_2) == "784") | (as.character( myData_d$diag_2) >= "790" & as.character( myData_d$diag_2) <= "799")
                    | (as.character( myData_d$diag_2) >= "240" & as.character( myData_d$diag_2) <= "249") | (as.character( myData_d$diag_2) >= "251" & as.character( myData_d$diag_2) <= "279")
                    | (as.character( myData_d$diag_2) >= "680" & as.character( myData_d$diag_2) <= "709") | (as.character( myData_d$diag_2) == "782") 
                    | (as.character( myData_d$diag_2) >= "001" & as.character( myData_d$diag_2) <= "139") | (as.character( myData_d$diag_2) >= "290" & as.character( myData_d$diag_2) <= "319")
                    | (as.character( myData_d$diag_2) >= "280" & as.character( myData_d$diag_2) <= "289") | (as.character( myData_d$diag_2) >= "320" & as.character( myData_d$diag_2) <= "359")
                    | (as.character( myData_d$diag_2) >= "630" & as.character( myData_d$diag_2) <= "679") | (as.character( myData_d$diag_2) >= "360" & as.character( myData_d$diag_2) <= "389")
                    | (as.character( myData_d$diag_2) >= "740" & as.character( myData_d$diag_2) <= "759")
                    | (startsWith(as.character( myData_d$diag_2), 'E')) 
                    | (startsWith(as.character( myData_d$diag_2), 'V'))
                    | (as.character( myData_d$diag_3) == "780") | (as.character( myData_d$diag_3) == "781")
                    | (as.character( myData_d$diag_3) == "784") | (as.character( myData_d$diag_3) >= "790" & as.character( myData_d$diag_3) <= "799")
                    | (as.character( myData_d$diag_3) >= "240" & as.character( myData_d$diag_3) <= "249") | (as.character( myData_d$diag_3) >= "251" & as.character( myData_d$diag_3) <= "279")
                    | (as.character( myData_d$diag_3) >= "680" & as.character( myData_d$diag_3) <= "709") | (as.character( myData_d$diag_3) == "782")
                    | (as.character( myData_d$diag_3) >= "001" & as.character( myData_d$diag_3) <= "139") | (as.character( myData_d$diag_3) >= "290" & as.character( myData_d$diag_3) <= "319")
                    | (as.character( myData_d$diag_3) >= "280" & as.character( myData_d$diag_3) <= "289") | (as.character( myData_d$diag_3) >= "320" & as.character( myData_d$diag_3) <= "359")
                    | (as.character( myData_d$diag_3) >= "630" & as.character( myData_d$diag_3) <= "679") | (as.character( myData_d$diag_3) >= "360" & as.character( myData_d$diag_3) <= "389")
                    | (as.character( myData_d$diag_3) >= "740" & as.character( myData_d$diag_3) <= "759")
                    | (startsWith(as.character( myData_d$diag_3), 'E')) 
                    | (startsWith(as.character( myData_d$diag_3), 'V'))] <- 1
```

```{r}
readmission <- myData_d
```

```{r}
head(readmission[, c(7:13, 17, 45:54)])
```
```{r}
readmission <- readmission[, c(7:13, 17, 45:54)]
```

```{r}
head(readmission)
```

```{r}
#combining dummy variables with numerical dataset
readmission<-cbind(readmission,d)
```

```{r}
head(readmission)
```
```{r}
nrow(readmission)
ncol(readmission)
```
```{r}
#change outcome variable to factor
readmission$readmitted<-as.factor(readmission$readmitted)
```

```{r}
class(readmission$readmitted)
```
```{r}
str(readmission)
```

```{r}
#rearrange dataset to have outcome variable in last index position
readmission<-readmission[,c(1,2,3,4,5,6,7,8,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,9)]
```

```{r}
head(readmission)
```

##Machine Learning-unsupervised learning

#k-means clustering
```{r}
#k-means clustering with numerical dataset
sapply(readmission_num,function(x)sum(is.na(x)))
```
```{r}
wcss2 = vector()
for (i in 1:10){
    model = kmeans(readmission_num, i)
    wcss2[i] = sum(model$withinss)
}
```

```{r}
wcss = vector()
for (i in 1:10){
    print(i)
    model = kmeans(readmission_num, i)
    wcss[i] = sum(model$withinss)
    print(wcss[i])
}
```
```{r}
wcss = vector()
for (i in 1:10){
    model = kmeans(readmission_num, i)
    wcss[i] = sum(model$withinss)
}

plot(1:10,
     wcss,
     type = 'b',
     main = paste('The Elbow Method'),
     xlab = 'Number of clusters',
     ylab = 'WCSS')
```
```{r}
kmeans = kmeans(x = readmission_num, centers = 3)
```
```{r}
kmeans$cluster
```
```{r}
kmeans$centers
```
```{r}
wcss = kmeans$withinss
wcss
```
```{r}
sum(kmeans$withinss)
```
```{r}
kmeans$tot.withinss
```
```{r}
y_kmeans = kmeans$cluster
```

```{r}
# install.packages('cluster')
```

```{r}
library(cluster)
```
```{r}
clusplot(readmission_num,
         y_kmeans,
         lines = 0,
         shade = TRUE,
         color = TRUE,
         labels = 2,
         plotchar = FALSE,
         span = TRUE,
         main = 'Clusters of readmissions',
         xlab = 'time in hospital',
         ylab = 'number of diagnoses')
```

#principal component analysis(pca)
```{r}
# install.packages('caret')
# install.packages('e1071')
library(caret)
library(e1071)
```

```{r}
readmission.pca <- prcomp(readmission_num, center = TRUE,scale. = TRUE)
```

```{r}
summary(readmission.pca)
```
```{r}
str(readmission.pca)
```

```{r}
library("factoextra")
get_eigenvalue(readmission.pca)
```

```{r}
#using variance:
library("factoextra")
fviz_eig(readmission.pca)
```
```{r}
# install.packages("remotes")
# remotes::install_github("vqv/ggbiplot")
```

```{r}
library(ggbiplot)
ggbiplot(readmission.pca)
```


##Machine Learning-supervised learning
```{r}
#split data into train and test
library(caTools)
set.seed(123)
split = sample.split(readmission$readmitted, SplitRatio = 0.7)
training_set = subset(readmission, split == TRUE)
test_set = subset(readmission, split == FALSE)
```

#logistic regression
```{r}
#logistic regression
log.model.1 <- glm(formula=readmitted ~ . , family=binomial,data = training_set)
```
```{r}
summary(log.model.1)
```


```{r}
#logistic regression
log.model.2 <- glm(formula=readmitted ~diag_diab+num_procedures+number_inpatient+A1Cresult8+admission_source_id.1 , family=binomial,data = training_set)
```
```{r}
summary(log.model.2)
```


```{r}
#variable importance for logistic regression
library(caret)
importance<-varImp(log.model.1, scale=FALSE)
importance
```
```{r}
#count number of NA's in test data
sum(is.na(test_set))
```
```{r}
results = predict(log.model.2, test_set, type='response')
results
```
```{r}
results_class <- ifelse(results>0.5, 1,0)
results_class
```
```{r}
table(test_set$readmitted, results_class)
```
```{r}
misClasificError <- mean(results_class != test_set$readmitted)
print(paste('Accuracy',1-misClasificError))
```
```{r}
#sensitivity=TP/TP+FN
sensitivity=30/(30+23)
sensitivity
#spec=TN/TN+FP
spec=22/(22+15)
spec
```


```{r}
library(pROC)
roc(test_set$readmitted, results, plot=TRUE,)
```
```{r}
auc(test_set$readmitted, results)
```


#random forest
```{r}
#random forest
#install.packages('randomForest')
library(randomForest)
```

```{r}
RF <- randomForest(formula = readmitted ~ ., method='class',
                 data = training_set)
print(RF)
```
```{r}
importance(RF)
```
```{r}
#MeanDecreaseGin --> Importance (IncNodePurity)
test_set[-42]
```

```{r}
# Predicting the Test set results
y_pred = predict(RF, newdata = test_set)
```

```{r}
confusionMatrix(test_set[, 42], y_pred)
```

#Xgboost
```{r}
#Xgboost
# install.packages('xgboost')
library(xgboost)
```
```{r}
classifier = xgboost(data = as.matrix(training_set[-42]), label = training_set$readmitted, nrounds = 10)
```
```{r}
y_pred = predict(classifier, newdata = as.matrix(test_set[-42]))
```

```{r}
head(y_pred)
```
```{r}
y_pred = (y_pred >= 0.5)
```

```{r}
head(y_pred)
```
```{r}
cm = table(test_set[, 42], y_pred)
cm
```
```{r}
imp_matrix = xgb.importance(model=classifier)
imp_matrix
xgb.plot.importance(imp_matrix)
```
```{r}
folds = createFolds(training_set$readmitted, k = 10)
class(folds)
```


```{r}
cv = lapply(folds, function(x) {
  training_fold = training_set[-x, ]
  test_fold = training_set[x, ]
  classifier = xgboost(data = as.matrix(training_fold[-42]), label = training_fold$readmitted, nrounds = 10)
  y_pred = predict(classifier, newdata = as.matrix(test_fold[-42]))
  y_pred = (y_pred >= 0.5)
  cm = table(test_fold[, 42], y_pred)
  accuracy = (cm[2,1]) / (cm[1,1] + cm[2,1])
  return(accuracy)
})
accuracy = mean(as.numeric(cv))
accuracy
```
```{r}
cv
```


```{r}
#feature scaling 
library(caret)
normParam <- preProcess(training_set, method = c("center", "scale"))
training_set <- predict(normParam, training_set)
test_set <- predict(normParam, test_set)
```

#KNN
```{r}
#KNN
library(class)
y_pred = knn(train = training_set[, -42],
             test = test_set[, -42],
             cl = training_set[, 42],
             k = 5)

y_pred
```

```{r}
# Fitting KNN to the Training set and Predicting the Test set results
library(class)
vec = c()
k_vec = c()
for (k in 1:50){
y_pred = knn(train = training_set[, -42],
             test = test_set[, -42],
             cl = training_set[, 42],
             k = 8)
y_pred

error = mean(y_pred != test_set$readmitted)
k_vec = c(k_vec, k)
vec = c(vec, error)}
```

```{r}
vec
```

```{r}
df.error = data.frame(k_vec,vec)
df.error
```

```{r}
ggplot(df.error, aes(k_vec, vec))+geom_line()
```

```{r}
y_pred = knn(train = training_set[, -42],
             test = test_set[, -42],
             cl = training_set[, 42],
             k = 18)
y_pred
# Making the Confusion Matrix
cm = table(test_set[, 42], y_pred)
cm
```

```{r}
confusionMatrix(test_set[, 42], y_pred)
```

#SVM
```{r}
# Fitting SVM to the Training set
library(e1071)
classifier = svm(formula = readmitted ~ .,
                 data = training_set,
                 kernel = 'linear')
```
```{r}
# Predicting the Test set results
y_pred = predict(classifier, newdata = test_set[-42])
print(y_pred)
# Making the Confusion Matrix
cm = table(test_set[, 42], y_pred)
cm
```
```{r}
confusionMatrix(test_set[, 42], y_pred)
```
```{r}
# Fitting SVM to the Training set with radial kernel
#install.packages('e1071')
library(e1071)
classifier = svm(formula = readmitted ~ .,
                 data = training_set,
                 kernel = 'radial')

# Predicting the Test set results
y_pred = predict(classifier, newdata = test_set[-42])
# 
# # Making the Confusion Matrix
cm = table(test_set[, 42], y_pred)
summary(classifier)
```
```{r}
confusionMatrix(test_set[, 42], y_pred)
```
```{r}
summary(classifier)
```
```{r}
#tuning
tune.results <- tune(svm,train.x=training_set[,-42],train.y=training_set[,42],kernel='radial',
                  ranges=list(cost=10^(-1:2), gamma=c(0.25,.5,1,2)))
```


```{r}
summary(tune.results)
```

```{r}
classifier = svm(formula = readmitted ~ .,
                 data = training_set,
                 type = 'C-classification',
                 kernel = 'radial',
                 cost=10,
                 gamma=0.25)

#Predicting the Test set results
y_pred = predict(classifier, newdata = test_set[-42])

#Making the Confusion Matrix
cm = table(test_set[, 42], y_pred)
cm
```
```{r}
confusionMatrix(test_set[, 42], y_pred)
```
```{r}
summary(classifier)
```

#Assessing the models-cross validation
```{r}
#cross-validation
train_control <- trainControl(method = "cv",
                              number = 5,
                              classProbs = TRUE,
                              summaryFunction=twoClassSummary)
model <- train(readmitted ~., data = readmission,
               method = "rf",
               metric="ROC",
               trControl = train_control)
```

```{r}
print(model)
```
```{r}
# install.packages('kernlab')
```

```{r}
train_control <- trainControl(method = "cv",
                              number = 5,
                              classProbs = TRUE,
                              summaryFunction=twoClassSummary)
model <- train(readmitted ~., data = readmission,
               method = "svmLinear",
               metric="ROC",
               trControl = train_control)
```
```{r}
print(model)
```

##Conclusion

The dataset had 101766 rows and 50 columns.I decided to remove the columns "encounter_id", "patient_nbr" because they are not important, and "weight", "payer_code", and "medical_specialty" since they had a lot of missing values.I also removed the rows with "None" value in the features "max_glu_serum" and "A1Cresult" as it would not provide useful information.I was left with 298 rows and 45 columns after cleaning.

After that, I classified the outcome variable into 2 groups:"YES" and "NO" using a criteria that would be more balanced.It would be difficult to encode dummy variables for "diag_1", "diag_2", and "diag_3" since they have a lot of levels, so I created diagnosis variables with ICD-9 codes.The ML models included k-means clustering, PCA,logistic regression, random forest,Xgboost, KNN, SVM,and cross-validation, using performance metrics such as accuracy, sensitivity,specificity, AUC, and AIC. From my analysis, the winning model is SVM with the highest accuracy of 0.64.

